---
title: "Hokkaidō Island: Data Exploration"
author: "Qiming Tian"
format: revealjs
execute:
  echo: false
  warning: false
  message: false
---

```{r}
library(dplyr)
library(readr)
library(ggplot2)
pop <- read_csv("../data/hokkaido_population.csv")
climate <- read_csv("../data/hokkaido_climate.csv")
eq <- read_csv("../data/earthquakes_japan.csv")
```

# Overview

- Population trends  
- Climate changes  
- Earthquake activity  
- Key insights  

---

# Hokkaidō

- Northernmost island of Japan  
- Cold climate with heavy snowfall  
- Known for agriculture, winter tourism, and earthquakes  

---

# Data Sources

- Japan population statistics  
- Climate records (monthly temperatures)  
- USGS-style earthquake catalog  

---

# Population Trend

```{r}
ggplot(pop, aes(year, population)) +
  geom_line() +
  theme_minimal()
```

---

# Aging Population

```{r}
ggplot(pop, aes(year, pct_over65)) +
  geom_line() +
  theme_minimal()
```

---

# Climate Warming

```{r}
climate_annual <- climate |>
  group_by(year) |>
  summarise(mean_temp = mean(mean_temp, na.rm = TRUE))

ggplot(climate_annual, aes(year, mean_temp)) +
  geom_line() +
  geom_smooth(se = FALSE) +
  theme_minimal()
```

---

# Seasonal Temperature

```{r}
ggplot(climate, aes(factor(month), mean_temp)) +
  geom_boxplot() +
  theme_minimal()
```

---

# Earthquake Distribution

```{r}
ggplot(eq, aes(mag)) +
  geom_histogram(bins = 30) +
  theme_minimal()
```

---

# Key Insights

- Persistent population decline and rapid aging  
- Clear long-term warming trend  
- Ongoing earthquake exposure  

---

# Thank You
